---
title: "biostats_project"
author: "Joe LaRocca"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(glmnet)
```

## Upload Data

```{r}

proj_1_data = read_csv("data/Project_1_Data.csv") |>
  janitor::clean_names() |>
  mutate(wkly_study_hours = 
           case_match(
             wkly_study_hours,
             "< 5" ~ "< 5 hours",
             "10-May" ~ "5-10 hours",
             "> 10" ~ "> 10 hours"
           ),
         ethnic_group = 
           case_match(
             ethnic_group,
             "group A" ~ "A",
             "group B" ~ "B",
             "group C" ~ "C",
             "group D" ~ "D",
             "group E" ~ "E"
           )
         ) |>
  drop_na()

```

## Create Table Summarizing Response Variables

```{r}

proj_1_data |>
  select(math_score, reading_score, writing_score) |>
  pivot_longer(
    cols= everything(),
    names_to = "Test",
    values_to = "Score"
  ) |>
  group_by(Test) |>
  summarize(Mean = round(mean(Score), 2),
            StDev = round(sd(Score), 2),
            Min = min(Score),
            Max = max(Score)) |>
  mutate(Test = case_match(
    Test,
    "math_score" ~ "Math",
    "reading_score" ~ "Reading",
    "writing_score" ~ "Writing"
    )
  ) |>
  knitr::kable()

```

## Single-Variable Categorical Plots

### Ethnic Group

```{r}

proj_1_data |>
  ggplot(aes(x = ethnic_group)) + 
  geom_bar()

```

### Parent Education

```{r}

proj_1_data |>
  ggplot(aes(x = parent_educ)) + 
  geom_bar()

```

### Lunch Type

```{r}

proj_1_data |>
  ggplot(aes(x = lunch_type)) + 
  geom_bar()

```

### Test Prep

```{r}

proj_1_data |>
  ggplot(aes(x = test_prep)) + 
  geom_bar()

```

### Parent Marital Status

```{r}

proj_1_data |>
  ggplot(aes(x = parent_marital_status)) + 
  geom_bar()

```

### Practicing Sports

```{r}

proj_1_data |>
  ggplot(aes(x = practice_sport)) + 
  geom_bar()

```

### First Child

```{r}

proj_1_data |>
  ggplot(aes(x = is_first_child)) + 
  geom_bar()

```

### Number of Siblings

```{r}

proj_1_data |>
  ggplot(aes(x = nr_siblings)) + 
  geom_bar()

```

# Modeling

### Create Data Frames with only Math, Reading, Writing to Make LM Functions Easier to Write

```{r}

math_data = proj_1_data |>
  select(-reading_score, -writing_score)

reading_data = proj_1_data |>
  select(-math_score, -writing_score)

writing_data = proj_1_data |>
  select(-math_score, -reading_score)

```

## All Predictors

### Math_All

```{r}

lm_all_math = lm(math_score ~ ., data = math_data)

summary(lm_all_math)

lm_all_math |> broom::glance() |> select(adj.r.squared)

```

### Reading_All

```{r}

lm_all_reading = lm(reading_score ~ ., data = reading_data)

summary(lm_all_reading)

lm_all_reading |> broom::glance() |> select(adj.r.squared)

```

### Writing_All

```{r}

lm_all_writing = lm(writing_score ~ ., data = writing_data)

summary(lm_all_writing)

lm_all_writing |> broom::glance() |> select(adj.r.squared)

```

## Backward selection

### Math
```{r}
math_back_lm <- step(lm_all_math, direction = "backward")

summary(math_back_lm)
```

### Reading
```{r}
read_back_lm <- step(lm_all_reading, direction = "backward")

summary(read_back_lm)
```

### Writing
```{r}
write_back_lm <- step(lm_all_writing, direction = "backward")

summary(write_back_lm)
```
