---
title: "biostats_project"
author: "Joe LaRocca"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(glmnet)
library(fastDummies)
```

## Upload Data

```{r}

proj_1_data = read_csv("data/Project_1_data.csv") |>
  janitor::clean_names() |>
  mutate(wkly_study_hours = 
           case_match(
             wkly_study_hours,
             "< 5" ~ "< 5 hours",
             "10-May" ~ "5-10 hours",
             "> 10" ~ "> 10 hours"
           ),
         ethnic_group = 
           case_match(
             ethnic_group,
             "group A" ~ "A",
             "group B" ~ "B",
             "group C" ~ "C",
             "group D" ~ "D",
             "group E" ~ "E"
           ),
         nr_siblings = 
           case_match(
             nr_siblings,
             0 ~ "0",
             1 ~ "1",
             2 ~ "2",
             3 ~ "3",
             4 ~ "4+",
             5 ~ "4+",
             6 ~ "4+",
             7 ~ "4+"
           )
         ) 
```


Filling NA's. Creating Model Matrix
```{r}

dummy_col_names = c(
  "gender_male" 
  #, "gender_female"
  
  # , "ethnic_group_A",
  , "ethnic_group_B", "ethnic_group_C" , "ethnic_group_D" , "ethnic_group_E"
  
  # , "parent_educ_some high school"
  , "parent_educ_high school", "parent_educ_some college"
  , "parent_educ_associate's degree", "parent_educ_bachelor's degree"
  , "parent_educ_master's degree" 
  
  
  # , "lunch_type_standard"
  ,"lunch_type_free/reduced"
  
  # , "test_prep_none"
  , "test_prep_completed"
  
  # ,  "parent_marital_status_married"
  , "parent_marital_status_divorced", "parent_marital_status_single"
  , "parent_marital_status_widowed"
  
  # , "practice_sport_never"
  , "practice_sport_sometimes", "practice_sport_regularly"
  
  # , "is_first_child_no"
  , "is_first_child_yes"
  
  # , "transport_means_school_bus"
  ,  "transport_means_private"
  
  # , "wkly_study_hours_< 5 hours"
  , "wkly_study_hours_> 10 hours", "wkly_study_hours_5-10 hours"
)



score_dummies =
  proj_1_data %>% 
  select(-math_score, -reading_score, -writing_score) %>% 
  dummy_cols(ignore_na = TRUE) %>% 
  select(all_of(dummy_col_names)) %>% 
  ## fill NA's
  lapply(function(x) ifelse(is.na(x), round(mean(x, na.rm = TRUE), 3), x)) %>% 
  as.data.frame()
```




```{r}
score_dummies
```



## Create Table Summarizing Response Variables

```{r}

proj_1_data |>
  select(math_score, reading_score, writing_score) |>
  pivot_longer(
    cols= everything(),
    names_to = "Test",
    values_to = "Score"
  ) |>
  group_by(Test) |>
  summarize(Mean = round(mean(Score), 2),
            StDev = round(sd(Score), 2),
            Min = min(Score),
            Max = max(Score)) |>
  mutate(Test = case_match(
    Test,
    "math_score" ~ "Math",
    "reading_score" ~ "Reading",
    "writing_score" ~ "Writing"
    )
  ) |>
  knitr::kable()

```

## Single-Variable Categorical Plots

### Ethnic Group

```{r}

proj_1_data |>
  ggplot(aes(x = ethnic_group)) + 
  geom_bar()

```

### Parent Education

```{r}

proj_1_data |>
  ggplot(aes(x = parent_educ)) + 
  geom_bar()

```

### Lunch Type

```{r}

proj_1_data |>
  ggplot(aes(x = lunch_type)) + 
  geom_bar()

```

### Test Prep

```{r}

proj_1_data |>
  ggplot(aes(x = test_prep)) + 
  geom_bar()

```

### Parent Marital Status

```{r}

proj_1_data |>
  ggplot(aes(x = parent_marital_status)) + 
  geom_bar()

```

### Practicing Sports

```{r}

proj_1_data |>
  ggplot(aes(x = practice_sport)) + 
  geom_bar()

```

### First Child

```{r}

proj_1_data |>
  ggplot(aes(x = is_first_child)) + 
  geom_bar()

```

### Number of Siblings

```{r}

proj_1_data |>
  ggplot(aes(x = nr_siblings)) + 
  geom_bar()

```

### Further EDA

```{r}
# Histograms for Response Variables

ggplot(data=proj_1_data, aes(math_score)) +
  geom_histogram(color = "black", fill = "white") +
  labs(title = "Histogram of Math Scores",
       x = "Math Score",
       y = "") +
  theme_bw()
ggplot(data=proj_1_data, aes(reading_score)) +
  geom_histogram(color = "black", fill = "white") +
  labs(title = "Histogram of Reading Scores",
       x = "Reading Score",
       y = "") +
  theme_bw()
ggplot(data=proj_1_data, aes(writing_score)) +
  geom_histogram(color = "black", fill = "white") +
  labs(title = "Histogram of Writing Scores",
       x = "Writing Score",
       y = "") +
  theme_bw()

# Boxplots for Predictive Variables

ggplot(data=proj_1_data, aes(x=parent_educ,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parent Education",
       x = "Parent Level of Education",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=parent_educ,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parent Education",
       x = "Parent Level of Education",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=parent_educ,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parent Education",
       x = "Parent Level of Education",
       y = "Writng Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=wkly_study_hours,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Study Hours",
       x = "Amount of Hours Studied",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=wkly_study_hours,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Study Hours",
       x = "Amount of Hours Studied",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=wkly_study_hours,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Study Hours",
       x = "Amount of Hours Studied",
       y = "Writing Score") +
  theme_bw()

# Looks strong
ggplot(data=proj_1_data, aes(x=lunch_type,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Lunch Type",
       x = "Lunch Type",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=lunch_type,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Lunch Type",
       x = "Lunch Type",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=lunch_type,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Lunch Type",
       x = "Lunch Type",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=transport_means,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Transportation",
       x = "Transportation Type",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=transport_means,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Transportation",
       x = "Transportation Type",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=transport_means,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Transportation",
       x = "Transportation Type",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=practice_sport,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Sports",
       x = "Sports Participation",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=practice_sport,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Sports",
       x = "Sports Participation",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=practice_sport,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Sports",
       x = "Sports Participation",
       y = "Writing Score") +
  theme_bw()

# Looks strong, look at diff between math and read/write
ggplot(data=proj_1_data, aes(x=gender,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Gender",
       x = "Gender",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=gender,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Gender",
       x = "Gender",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=gender,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Gender",
       x = "Gender",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=ethnic_group,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Ethnic Group",
       x = "Ethnic Group",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=ethnic_group,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Ethnic Group",
       x = "Ethnic Group",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=ethnic_group,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Ethnic Group",
       x = "Ethnic Group",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=parent_marital_status,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parental Marital Status",
       x = "Parental Marital Status",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=parent_marital_status,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parental Marital Status",
       x = "Parental Marital Status",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=parent_marital_status,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parental Marital Status",
       x = "Parental Marital Status",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=is_first_child,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for First Child Status",
       x = "First Child Status",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=is_first_child,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for First Child Status",
       x = "First Child Status",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=is_first_child,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for First Child Status",
       x = "First Child Status",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=nr_siblings,y=math_score)) + geom_point()
ggplot(data=proj_1_data, aes(x=nr_siblings,y=reading_score)) + geom_point()
ggplot(data=proj_1_data, aes(x=nr_siblings,y=writing_score)) + geom_point()
```

# Modeling

### Create Data Frames with only Math, Reading, Writing to Make LM Functions Easier to Write

```{r}

math_data = proj_1_data |>
  select(-reading_score, -writing_score)

reading_data = proj_1_data |>
  select(-math_score, -writing_score)

writing_data = proj_1_data |>
  select(-math_score, -reading_score)

```

## All Predictors

### Math_All

```{r}

lm_all_math = lm(math_score ~ ., data = math_data)

summary(lm_all_math)

lm_all_math |> broom::glance() |> select(adj.r.squared)

```

### Reading_All

```{r}

lm_all_reading = lm(reading_score ~ ., data = reading_data)

summary(lm_all_reading)

lm_all_reading |> broom::glance() |> select(adj.r.squared)

```

### Writing_All

```{r}

lm_all_writing = lm(writing_score ~ ., data = writing_data)

summary(lm_all_writing)

lm_all_writing |> broom::glance() |> select(adj.r.squared)

```

## LASSO

### Math_LASSO

```{r}

math_pred = model.matrix(math_score ~ ., math_data)[, -1]

math_score = pull(math_data, math_score)

math_cv = cv.glmnet(x = math_pred, y = math_score)

math_lasso = glmnet(math_pred, math_score, lambda = math_cv$lambda.min) 

r2_math_lasso = math_lasso$dev.ratio

adj_r2_math_lasso = 1 - ((1 - r2_math_lasso) * (nrow(math_data) - 1)) / (nrow(math_data) - nrow(broom::tidy(math_lasso)) - 1)

adj_r2_math_lasso

```

### Reading_LASSO

```{r}

reading_pred = model.matrix(reading_score ~ ., reading_data)[, -1]

reading_score = pull(reading_data, reading_score)

reading_cv = cv.glmnet(x = reading_pred, y = reading_score)

reading_lasso = glmnet(reading_pred, reading_score, lambda = reading_cv$lambda.min) 

r2_reading_lasso = reading_lasso$dev.ratio

adj_r2_reading_lasso = 1 - ((1 - r2_reading_lasso) * (nrow(reading_data) - 1)) / (nrow(reading_data) - nrow(broom::tidy(reading_lasso)) - 1)

adj_r2_reading_lasso

```

### Writing_LASSO

```{r}

writing_pred = model.matrix(writing_score ~ ., writing_data)[, -1]

writing_score = pull(writing_data, writing_score)

writing_cv = cv.glmnet(x = writing_pred, y = writing_score)

writing_lasso = glmnet(writing_pred, writing_score, lambda = writing_cv$lambda.min) 

r2_writing_lasso = writing_lasso$dev.ratio

adj_r2_writing_lasso = 1 - ((1 - r2_writing_lasso) * (nrow(writing_data) - 1)) / (nrow(writing_data) - nrow(broom::tidy(writing_lasso)) - 1)

adj_r2_writing_lasso

```

