---
title: "Biostatistical Methods I - Final Project Code"
author: "Ravi Brenner, Cameron Chesbrough, Joe LaRocca, Wayne Monical"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(glmnet)
library(fastDummies)
library(MASS)
library(tidyverse)

```

## Upload Data

```{r}

proj_1_data = read_csv("data/Project_1_data.csv") |>
  janitor::clean_names() |>
  mutate(wkly_study_hours = 
           case_match(
             wkly_study_hours,
             "< 5" ~ "< 5 hours",
             "10-May" ~ "5-10 hours",
             "> 10" ~ "> 10 hours"
           ),
         ethnic_group = 
           case_match(
             ethnic_group,
             "group A" ~ "A",
             "group B" ~ "B",
             "group C" ~ "C",
             "group D" ~ "D",
             "group E" ~ "E"
           ),
         nr_siblings = 
           case_match(
             nr_siblings,
             0 ~ "0",
             1 ~ "1",
             2 ~ "2",
             3 ~ "3",
             4 ~ "4+",
             5 ~ "4+",
             6 ~ "4+",
             7 ~ "4+"
           )
         ) 
```


Filling NA's. Creating Model Matrix
```{r}

dummy_col_names = c(
  "gender_male" 
  #, "gender_female"
  
  # , "ethnic_group_A",
  , "ethnic_group_B", "ethnic_group_C" , "ethnic_group_D" , "ethnic_group_E"
  
  # , "parent_educ_some high school"
  , "parent_educ_high school", "parent_educ_some college"
  , "parent_educ_associate's degree", "parent_educ_bachelor's degree"
  , "parent_educ_master's degree" 
  
  
  # , "lunch_type_standard"
  ,"lunch_type_free/reduced"
  
  # , "test_prep_none"
  , "test_prep_completed"
  
  # ,  "parent_marital_status_married"
  , "parent_marital_status_divorced", "parent_marital_status_single"
  , "parent_marital_status_widowed"
  
  # , "practice_sport_never"
  , "practice_sport_sometimes", "practice_sport_regularly"
  
  # , "is_first_child_no"
  , "is_first_child_yes"
  
  # , "transport_means_school_bus"
  ,  "transport_means_private"
  
  # , "nr_siblings_0"
  , "nr_siblings_1", "nr_siblings_2", "nr_siblings_3", "nr_siblings_4+" 
  
  # , "wkly_study_hours_< 5 hours"
  , "wkly_study_hours_> 10 hours", "wkly_study_hours_5-10 hours"
)

score_dummies =
  proj_1_data %>% 
  dplyr::select(-math_score, -reading_score, -writing_score) %>% 
  dummy_cols(ignore_na = TRUE) %>% 
  dplyr::select(all_of(dummy_col_names)) %>% 
  ## fill NA's
  lapply(function(x) ifelse(is.na(x), round(mean(x, na.rm = TRUE), 3), x)) %>% 
  as.data.frame()
```



### Correlation

```{r}
cor_df = 
  proj_1_data %>%
  dplyr::select(-reading_score, -writing_score) %>% 
  drop_na()

cor_df = 
  model.matrix(object = math_score ~ ., data = cor_df) %>% 
  cor() %>% 
  as.data.frame()

cor_df = 
  cor_df %>% 
  mutate(x_name = row.names(cor_df)) %>% 
  filter(x_name != '(Intercept)') %>% 
  dplyr::select(-'(Intercept)') %>% 
  pivot_longer(
    gendermale:`wkly_study_hours5-10 hours`
    , values_to = 'correlation'
    , names_to = 'y_name'
  ) %>% 
  filter(
    substr(x_name, 1, 6) != substr(y_name, 1, 6),
    tolower(x_name) < tolower(y_name)
  )
```

Getting a histogram of the correlations between the variables.
```{r}
hist(cor_df$correlation,
     main = "Histogram of Correlations Between Predictors",
     xlab = "Correlation Coefficient")
```

Getting the top correlations
```{r}
cor_df %>% 
  arrange(desc(correlation)) %>% 
  mutate(correlation = round(correlation, digits = 3)) %>% 
  head(6) %>% 
  knitr::kable()
```





## Create Table Summarizing Response Variables

```{r}

proj_1_data |>
  dplyr::select(math_score, reading_score, writing_score) |>
  pivot_longer(
    cols= everything(),
    names_to = "Test",
    values_to = "Score"
  ) |>
  group_by(Test) |>
  summarize(Mean = round(mean(Score), 2),
            StDev = round(sd(Score), 2),
            Min = min(Score),
            Max = max(Score)) |>
  mutate(Test = case_match(
    Test,
    "math_score" ~ "Math",
    "reading_score" ~ "Reading",
    "writing_score" ~ "Writing"
    )
  ) |>
  knitr::kable()

```

## Single-Variable Categorical Plots

### Ethnic Group

```{r}

proj_1_data |>
  ggplot(aes(x = ethnic_group)) + 
  geom_bar()

```

### Parent Education

```{r}

proj_1_data |>
  ggplot(aes(x = parent_educ)) + 
  geom_bar()

```

### Lunch Type

```{r}

proj_1_data |>
  ggplot(aes(x = lunch_type)) + 
  geom_bar()

```

### Test Prep

```{r}

proj_1_data |>
  ggplot(aes(x = test_prep)) + 
  geom_bar()

```

### Parent Marital Status

```{r}

proj_1_data |>
  ggplot(aes(x = parent_marital_status)) + 
  geom_bar()

```

### Practicing Sports

```{r}

proj_1_data |>
  ggplot(aes(x = practice_sport)) + 
  geom_bar()

```

### First Child

```{r}

proj_1_data |>
  ggplot(aes(x = is_first_child)) + 
  geom_bar()

```

### Number of Siblings

```{r}

proj_1_data |>
  ggplot(aes(x = nr_siblings)) + 
  geom_bar()

```

### Further EDA

```{r}
# Histograms for Response Variables

ggplot(data=proj_1_data, aes(math_score)) +
  geom_histogram(color = "black", fill = "white") +
  labs(title = "Histogram of Math Scores",
       x = "Math Score",
       y = "") +
  theme_bw()
ggplot(data=proj_1_data, aes(reading_score)) +
  geom_histogram(color = "black", fill = "white") +
  labs(title = "Histogram of Reading Scores",
       x = "Reading Score",
       y = "") +
  theme_bw()
ggplot(data=proj_1_data, aes(writing_score)) +
  geom_histogram(color = "black", fill = "white") +
  labs(title = "Histogram of Writing Scores",
       x = "Writing Score",
       y = "") +
  theme_bw()

# Boxplots for Predictive Variables

ggplot(data=proj_1_data, aes(x=parent_educ,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parent Education",
       x = "Parent Level of Education",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=parent_educ,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parent Education",
       x = "Parent Level of Education",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=parent_educ,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parent Education",
       x = "Parent Level of Education",
       y = "Writng Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=wkly_study_hours,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Study Hours",
       x = "Amount of Hours Studied",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=wkly_study_hours,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Study Hours",
       x = "Amount of Hours Studied",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=wkly_study_hours,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Study Hours",
       x = "Amount of Hours Studied",
       y = "Writing Score") +
  theme_bw()

# Looks strong
ggplot(data=proj_1_data, aes(x=lunch_type,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Lunch Type",
       x = "Lunch Type",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=lunch_type,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Lunch Type",
       x = "Lunch Type",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=lunch_type,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Lunch Type",
       x = "Lunch Type",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=transport_means,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Transportation",
       x = "Transportation Type",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=transport_means,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Transportation",
       x = "Transportation Type",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=transport_means,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Transportation",
       x = "Transportation Type",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=practice_sport,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Sports",
       x = "Sports Participation",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=practice_sport,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Sports",
       x = "Sports Participation",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=practice_sport,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Sports",
       x = "Sports Participation",
       y = "Writing Score") +
  theme_bw()

# Looks strong, look at diff between math and read/write
ggplot(data=proj_1_data, aes(x=gender,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Gender",
       x = "Gender",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=gender,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Gender",
       x = "Gender",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=gender,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Gender",
       x = "Gender",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=ethnic_group,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Ethnic Group",
       x = "Ethnic Group",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=ethnic_group,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Ethnic Group",
       x = "Ethnic Group",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=ethnic_group,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Ethnic Group",
       x = "Ethnic Group",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=parent_marital_status,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parental Marital Status",
       x = "Parental Marital Status",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=parent_marital_status,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parental Marital Status",
       x = "Parental Marital Status",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=parent_marital_status,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for Parental Marital Status",
       x = "Parental Marital Status",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=is_first_child,y=math_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for First Child Status",
       x = "First Child Status",
       y = "Math Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=is_first_child,y=reading_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for First Child Status",
       x = "First Child Status",
       y = "Reading Score") +
  theme_bw()
ggplot(data=proj_1_data, aes(x=is_first_child,y=writing_score)) +
  geom_boxplot() +
  labs(title = "Boxplots for First Child Status",
       x = "First Child Status",
       y = "Writing Score") +
  theme_bw()

ggplot(data=proj_1_data, aes(x=nr_siblings,y=math_score)) + geom_point()
ggplot(data=proj_1_data, aes(x=nr_siblings,y=reading_score)) + geom_point()
ggplot(data=proj_1_data, aes(x=nr_siblings,y=writing_score)) + geom_point()
```

# Modeling

### Create Data Frames with only Math, Reading, Writing to Make LM Functions Easier to Write

```{r}

math_data = score_dummies |> 
  mutate(math_score = proj_1_data$math_score)

reading_data = score_dummies |>
  mutate(reading_score = proj_1_data$reading_score)

writing_data = score_dummies |>
  mutate(writing_score = proj_1_data$writing_score)

```

## All Predictors

### Math_All

```{r}

lm_all_math = lm(math_score ~ ., data = math_data)

summary(lm_all_math)

aic_all_math = AIC(lm_all_math)
bic_all_math = BIC(lm_all_math)
adjr2_all_math = lm_all_math |> broom::glance() |> dplyr::select(adj.r.squared)

```

### Reading_All

```{r}

lm_all_reading = lm(reading_score ~ ., data = reading_data)

summary(lm_all_reading)

aic_all_reading = AIC(lm_all_reading)
bic_all_reading = BIC(lm_all_reading)
adjr2_all_reading = lm_all_reading |> broom::glance() |> dplyr::select(adj.r.squared)

```

### Writing_All

```{r}

lm_all_writing = lm(writing_score ~ ., data = writing_data)

summary(lm_all_writing)

aic_all_writing = AIC(lm_all_writing)
bic_all_writing = BIC(lm_all_writing)
adjr2_all_writing = lm_all_writing |> broom::glance() |> dplyr::select(adj.r.squared)

```

## Economic Model

### Math_Econ

```{r}

lm_econ_math = lm(math_score ~ parent_educ_high.school + 
                    parent_educ_some.college + 
                    parent_educ_associate.s.degree + 
                    parent_educ_bachelor.s.degree + 
                    parent_educ_master.s.degree + 
                    lunch_type_free.reduced + 
                    test_prep_completed + transport_means_private, 
                  data = math_data)

summary(lm_econ_math)

aic_econ_math = AIC(lm_econ_math)
bic_econ_math = BIC(lm_econ_math)
adjr2_econ_math = lm_econ_math |> broom::glance() |> dplyr::select(adj.r.squared)

```

### Reading_Econ

```{r}

lm_econ_reading = lm(reading_score ~ parent_educ_high.school + 
                    parent_educ_some.college + 
                    parent_educ_associate.s.degree + 
                    parent_educ_bachelor.s.degree + 
                    parent_educ_master.s.degree + 
                    lunch_type_free.reduced + 
                    test_prep_completed + transport_means_private, 
                  data = reading_data)

summary(lm_econ_reading)

aic_econ_reading = AIC(lm_econ_reading)
bic_econ_reading = BIC(lm_econ_reading)
adjr2_econ_reading = lm_econ_reading |> broom::glance() |> dplyr::select(adj.r.squared)

```

### Writing_Econ

```{r}

lm_econ_writing = lm(writing_score ~ parent_educ_high.school + 
                    parent_educ_some.college + 
                    parent_educ_associate.s.degree + 
                    parent_educ_bachelor.s.degree + 
                    parent_educ_master.s.degree + 
                    lunch_type_free.reduced + 
                    test_prep_completed + transport_means_private, 
                  data = writing_data)

summary(lm_econ_writing)

aic_econ_writing = AIC(lm_econ_writing)
bic_econ_writing = BIC(lm_econ_writing)
adjr2_econ_writing = lm_econ_writing |> broom::glance() |> dplyr::select(adj.r.squared)

```

## Hard Worker Model

### Math_HW

```{r}

lm_hw_math = lm(math_score ~ wkly_study_hours_..10.hours + 
                  wkly_study_hours_5.10.hours + 
                  practice_sport_sometimes + 
                  practice_sport_regularly,
                  data = math_data)

summary(lm_hw_math)

aic_hw_math = AIC(lm_hw_math)
bic_hw_math = BIC(lm_hw_math)
adjr2_hw_math = lm_hw_math |> broom::glance() |> dplyr::select(adj.r.squared)

```

### Reading_HW

```{r}

lm_hw_reading = lm(reading_score ~ wkly_study_hours_..10.hours + 
                  wkly_study_hours_5.10.hours + 
                  practice_sport_sometimes + 
                  practice_sport_regularly,
                  data = reading_data)

summary(lm_hw_reading)

aic_hw_reading = AIC(lm_hw_reading)
bic_hw_reading = BIC(lm_hw_reading)
adjr2_hw_reading = lm_hw_reading |> broom::glance() |> dplyr::select(adj.r.squared)

```

### Writing_HW

```{r}

lm_hw_writing = lm(writing_score ~ wkly_study_hours_..10.hours + 
                  wkly_study_hours_5.10.hours + 
                  practice_sport_sometimes + 
                  practice_sport_regularly,
                  data = writing_data)

summary(lm_hw_writing)

aic_hw_writing = AIC(lm_hw_writing)
bic_hw_writing = BIC(lm_hw_writing)
adjr2_hw_writing = lm_hw_writing |> broom::glance() |> dplyr::select(adj.r.squared)

```

## Demographic Model

### Math_Demo

```{r}

lm_demo_math = lm(math_score ~ gender_male + 
                    ethnic_group_B +
                    ethnic_group_C +
                    ethnic_group_D + 
                    ethnic_group_E, data = math_data)

summary(lm_demo_math)

aic_demo_math = AIC(lm_demo_math)
bic_demo_math = BIC(lm_demo_math)
adjr2_demo_math = lm_demo_math |> broom::glance() |> dplyr::select(adj.r.squared)

```

### Reading_Demo

```{r}

lm_demo_reading = lm(reading_score ~ gender_male + 
                    ethnic_group_B +
                    ethnic_group_C +
                    ethnic_group_D + 
                    ethnic_group_E, data = reading_data)

summary(lm_demo_reading)

aic_demo_reading = AIC(lm_demo_reading)
bic_demo_reading = BIC(lm_demo_reading)
adjr2_demo_reading = lm_demo_reading |> broom::glance() |> dplyr::select(adj.r.squared)

```

### Writing_Demo

```{r}

lm_demo_writing = lm(writing_score ~ gender_male + 
                    ethnic_group_B +
                    ethnic_group_C +
                    ethnic_group_D + 
                    ethnic_group_E, data = writing_data)

summary(lm_demo_writing)

aic_demo_writing = AIC(lm_demo_writing)
bic_demo_writing = BIC(lm_demo_writing)
adjr2_demo_writing = lm_demo_writing |> broom::glance() |> dplyr::select(adj.r.squared)

```

## Family Structure Model

### Math_Fam

```{r}

lm_fam_math = lm(math_score ~ parent_marital_status_divorced +
                   parent_marital_status_single + 
                   parent_marital_status_widowed + 
                   is_first_child_yes + 
                   nr_siblings_1 + 
                   nr_siblings_2 + 
                   nr_siblings_3 + 
                   nr_siblings_4., 
                 data = math_data)

summary(lm_fam_math)

aic_fam_math = AIC(lm_fam_math)
bic_fam_math = BIC(lm_fam_math)
adjr2_fam_math = lm_fam_math |> broom::glance() |> dplyr::select(adj.r.squared)

```

### Reading_Fam

```{r}

lm_fam_reading = lm(reading_score ~ parent_marital_status_divorced +
                   parent_marital_status_single + 
                   parent_marital_status_widowed + 
                   is_first_child_yes + 
                   nr_siblings_1 + 
                   nr_siblings_2 + 
                   nr_siblings_3 + 
                   nr_siblings_4., 
                 data = reading_data)

summary(lm_fam_reading)

aic_fam_reading = AIC(lm_fam_reading)
bic_fam_reading = BIC(lm_fam_reading)
adjr2_fam_reading = lm_fam_reading |> broom::glance() |> dplyr::select(adj.r.squared)

```

### Writing_Fam

```{r}

lm_fam_writing = lm(writing_score ~ parent_marital_status_divorced +
                   parent_marital_status_single + 
                   parent_marital_status_widowed + 
                   is_first_child_yes + 
                   nr_siblings_1 + 
                   nr_siblings_2 + 
                   nr_siblings_3 + 
                   nr_siblings_4., 
                 data = writing_data)

summary(lm_fam_writing)

aic_fam_writing = AIC(lm_fam_writing)
bic_fam_writing = BIC(lm_fam_writing)
adjr2_fam_writing = lm_fam_writing |> broom::glance() |> dplyr::select(adj.r.squared)

```

## Backwards Selection

### Math

```{r}

lm_step_math = stepAIC(lm_all_math, direction = "backward")

aic_step_math = AIC(lm_step_math)
bic_step_math = BIC(lm_step_math)
adjr2_step_math = lm_step_math |> broom::glance() |> pull(adj.r.squared)

```

### Reading

```{r}

lm_step_reading = stepAIC(lm_all_reading, direction = "backward")

aic_step_reading = AIC(lm_step_reading)
bic_step_reading = BIC(lm_step_reading)
adjr2_step_reading = lm_step_reading |> broom::glance() |> pull(adj.r.squared)

```

### Writing

```{r}

lm_step_writing = stepAIC(lm_all_writing, direction = "backward")

aic_step_writing = AIC(lm_step_writing)
bic_step_writing = BIC(lm_step_writing)
adjr2_step_writing = lm_step_writing |> broom::glance() |> pull(adj.r.squared)

```

## Readable Tables of Results

### Math

```{r}

model_stats_math = tibble(
  math_model_type = c("All", "Backward Selection", "Economic", "Hard Worker", "Demographic", "Family Structure"),
  ADJ_R2 = c(adjr2_all_math,
             adjr2_step_math,
             adjr2_econ_math, 
             adjr2_hw_math,
             adjr2_demo_math,
             adjr2_fam_math),
  AIC = c(aic_all_math,
          aic_step_math,
          aic_econ_math,
          aic_hw_math,
          aic_demo_math,
          aic_fam_math),
  BIC = c(bic_all_math,
          bic_step_math,
          bic_econ_math,
          bic_hw_math,
          bic_demo_math,
          bic_fam_math)
)

model_stats_math |> knitr::kable()

```

### Reading

```{r}

model_stats_reading = tibble(
  reading_model_type = c("All", "Backward Selection", "Economic", "Hard Worker", "Demographic", "Family Structure"),
  ADJ_R2 = c(adjr2_all_reading, 
             adjr2_step_reading,
             adjr2_econ_reading, 
             adjr2_hw_reading,
             adjr2_demo_reading,
             adjr2_fam_reading),
  AIC = c(aic_all_reading,
          aic_step_reading,
          aic_econ_reading,
          aic_hw_reading,
          aic_demo_reading,
          aic_fam_reading),
  BIC = c(bic_all_reading,
          bic_step_reading,
          bic_econ_reading,
          bic_hw_reading,
          bic_demo_reading,
          bic_fam_reading)
)

model_stats_reading |> knitr::kable()

```

### Writing

```{r}

model_stats_writing = tibble(
  writing_model_type = c("All", "Backward Selection", "Economic", "Hard Worker", "Demographic", "Family Structure"),
  ADJ_R2 = c(adjr2_all_writing, 
             adjr2_step_writing,
             adjr2_econ_writing, 
             adjr2_hw_writing,
             adjr2_demo_writing,
             adjr2_fam_writing),
  AIC = c(aic_all_writing,
          aic_step_writing,
          aic_econ_writing,
          aic_hw_writing,
          aic_demo_writing,
          aic_fam_writing),
  BIC = c(bic_all_writing,
          bic_step_writing,
          bic_econ_writing,
          bic_hw_writing,
          bic_demo_writing,
          bic_fam_writing)
)

model_stats_writing |> knitr::kable()

```

Final model statistics
```{r}
t1 <- broom::tidy(lm_step_math) |> dplyr::select(-std.error)

t2 <- broom::tidy(lm_step_reading) |> dplyr::select(-std.error)

t3 <- broom::tidy(lm_step_writing) |> dplyr::select(-std.error)

```




# Model Diagnostics

Plotting the residuals against the fitted values, we see no significant relationship.
```{r}
plot(predict(lm_all_math, score_dummies), lm_all_math$residuals)
```

Examining the residuals for normality, we see some deviation, but not egregious deviation. 
```{r}
qqnorm(lm_all_math$residuals)
```

Examining the residuals versus leverage, we find that no point is exerting fully undue leverage. 

```{r}
plot(lm_all_math, which = 5)
```