---
title: "biostats_project"
author: "Joe LaRocca"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(glmnet)
```

## Upload Data


```{r}
score = read_csv("data/Project_1_data.csv") |>
  janitor::clean_names() |>
  mutate(wkly_study_hours = 
           case_match(
             wkly_study_hours,
             "< 5" ~ "< 5 hours",
             "10-May" ~ "5-10 hours",
             "> 10" ~ "> 10 hours"
           ),
         ethnic_group = 
           case_match(
             ethnic_group,
             "group A" ~ "A",
             "group B" ~ "B",
             "group C" ~ "C",
             "group D" ~ "D",
             "group E" ~ "E"
           )
         )

```


### Look at Correlation

```{r}

```



## Data Preparation

Create model matrix with one hot-encoding
```{r}
library(fastDummies)
score_dummies <- 
  score %>% 
  select(-math_score, -reading_score, -writing_score) %>% 
  dummy_cols(ignore_na = TRUE) %>% 
  select()

```



Impute missing values with the column average.
```{r}

```

## Data exploration

Create table of descriptive statistics.
```{r}

```



Visualization
```{r}

```


Search for outliers. Outliers are defined as points falling 1.5 * IQR outside Q1 or Q3
```{r}
# look for outliers in 

  # of siblings

  # reading

  # writing

  # math


# these are the only 
```


Search for co-linearity/high correlation in predictors
```{r}
score %>% 
  #select(-math_score, -reading_score, -writing_score) %>% 
  as.matrix(math_score~., data =_) %>% 
  cor()
```


Using VIF
Testing suite (specify ahead of time)
Optimal Models
MLR Full Model
Evaluate with adj R Square
Backwards AIC Model
Consider all 2 way interactions
LASSO
Consider all covariates and 2 and 3 way interactions. Conduct CV to find optimal lambda
Table with all of the models
Adj R-square
AIC



Full model
Backwards AIC model
LASSO
Interesting model 1
Interesting model 2
Reading
0.7
0.64
etc.




Writing
0.75
0.96






Math
0.8
.77









Interesting Models


Model Diagnostics
Consider interactions
Plotting
X-values versus residuals
Check for homoscedasticity
Qq-plot of residuals to check that they are normal
If the residuals are not normal, we will use the Box-Cox transformation to find the best . 
Residuals versus leverage
Detect any points that are exerting lots of leverage. We will remove any points that have ___ too much leverage. 
Adjusted R-Squared
CV
For all automatically selecting models, we will do a 10-fold cross validated MSE test. 
Interesting Models
These models are for the discussion section. We’ll include discussion of models that we have room for. We will not be rigorously validating these models. 
Economic Model: Using any proxy for economic status
ParentEduc
LunchType
TransportMeans
TestPrep
Hard Worker Model
WklyStudyHours
PracticeSport 
Family Structure Model
ParentMaritalStatus
NrSiblings
IsFirstChild
Demographics Model
Gender
EthnicGroup



